Tool for Visual Cluster Analysis and Consensus Clustering

### 2020-04-13 presentation/presentation.tex
presentation
presentation proto

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[font=small]{caption}
\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usepackage{animate}
\graphicspath{ {../thesis/img/} }

@@

Idea:
	\begin{itemize}
		\item Combine group of robust solutions into one
	\end{itemize}

\begin{frame}{Idea of our Tool: Ease of Use}
	\begin{itemize}
		\item bla
	\end{itemize}

@@

@@

	\begin{figure}
	  \centering
	    \includegraphics[width=.6\textwidth]{unob}
	  \caption{Data-View of the Tool}
	  \label{fig:data-view}
	\end{figure}

@@

	\begin{figure}
	  \centering
	    \includegraphics[width=.7\textwidth]{workflow-view-tasks}
	  \caption{Workflow-View of the Tool}
	  \label{fig:workflow-view}
	\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{meta-view}
  \caption{Meta-View of the Tool}
  \label{fig:meta-view}
\end{figure}

@@

\begin{frame}{Implementation}

	\begin{itemize}
		\item bla
	\end{itemize}

\begin{frame}{Tests}

	\begin{itemize}
		\item bla
	\end{itemize}

\begin{frame}{Future Work}

\begin{itemize}
  \item bla
\end{itemize}

@@

\begin{itemize}
  \item bla
\end{itemize}

@@

\begin{frame}{References}
\printbibliography
\end{frame}

>>>>

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[font=small]{caption}
\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{animate}
\graphicspath{ {../thesis/img/} }

@@

Idea:
	\begin{itemize}
		\item Combine group of robust solutions into one
	\end{itemize}
\end{frame}

@@

Aim: Facilitating use through clear separation

@@

	\begin{figure}
	  \centering
	    \includegraphics[width=.64\textwidth]{unob}
	  \caption{Data-View}
	  \label{fig:data-view}
	\end{figure}

\begin{frame}{The Tool: Data-View - Scatter Plot Matrix}

	\begin{figure}
	  \centering
	    \includegraphics[width=.97\textwidth]{scatter_matrix}
	  \caption{Scatter Plot Matrix}
	  \label{fig:scatter_matrix}
	\end{figure}

@@

	\begin{figure}
	  \centering
	    \includegraphics[width=.7\textwidth]{workflow-view-tasks}
	  \caption{Workflow-View}
	  \label{fig:workflow-view}
	\end{figure}

@@

	\begin{figure}
	  \centering
	    \includegraphics[width=.9\textwidth]{meta-view}
	  \caption{Meta-View}
	  \label{fig:meta-view}
	\end{figure}

@@

\begin{frame}{Implementation}
Used tools:
	\begin{itemize}
		\item Java 1.8, utilizing Streams for parallelization
		\item Libraries:
		\begin{itemize}
			\item  ELKI \cite{10.1007/978-3-540-69497-7_41} - Clustering
			\item WEKA \cite{10.1145/1656274.1656278} - IO
			\item Java Smile \cite{javasmile} - Additional Methods
		\end{itemize}
		\item Swing's JComponents and overriding the \emph{draw()} method
		\item[]
	\end{itemize}
Ease of extension:
	\begin{itemize}
		\item All selectable methods provide simple interfaces
	\end{itemize}
\end{frame}

\begin{frame}{Tests: Introduction}
We want to show that with our tool we can:

	\begin{itemize}
		\item Produce solutions better than any individual clustering result
		\item Obtain solutions unobtainable by single methods
		\item Find multiple alternative solutions which can be analyzed to find a fitting choice
		\item[]
	\end{itemize}

And do so in a straightforward and useful way:
	\begin{itemize}
		\item Letting a user test our tool
		\item Also showing real world test data-sets
	\end{itemize}
\end{frame}

\begin{frame}{Tests: Better than individual Solutions}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{better_base}
		\caption{Synthetic Data with Ground Truth}
		\label{fig:better_base}
	\end{figure}
\end{frame}

\begin{frame}{Tests: Better than individual Solutions}
Best individual result when sampling Lloyd's k-Means algorithm with \newline $k=2...20$ and $6$ samples per $k$:
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{better_base_best}
		\caption{Result of best k-Means run for example Data-Set}
		\label{fig:better_base_best}
	\end{figure}
Combining all solutions finds the ground truth exactly (without defining $k$)
\end{frame}

\begin{frame}{Tests: Unobtainable Solutions}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{unob}
		\caption{Synthetic Data with Ground Truth}
		\label{fig:better_base}
	\end{figure}
\end{frame}

\begin{frame}{Tests: Unobtainable Solutions}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.4\textwidth]{unob_wf}
		\label{fig:unob_wf}
	\end{figure}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{unob_optics_stable}
		\caption{OPTICS reachability plot}
		\label{fig:unob_optics_stable}
	\end{figure}
\end{frame}

\begin{frame}{Tests: Unobtainable Solutions}
	\begin{figure}[h]
	\centering
	\begin{subfigure}[t]{.32\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{unob_k3}
		  \caption{K-Means with $k=3$}
		  \label{fig:unob_k3}
	\end{subfigure}
	\begin{subfigure}[t]{.32\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{unob_k4ex}
		  \caption{K-Means with $k=4$}
		  \label{fig:unob_k4ex}
	\end{subfigure}
	\begin{subfigure}[t]{.32\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{unob_k5ex}
		  \caption{K-Means with $k=5$}
		  \label{fig:unob_k5ex}
	\end{subfigure}
	\begin{subfigure}[t]{.32\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{unob_best}
		  \caption{DBSCAN, best single Result}
		  \label{fig:unob_best}
	\end{subfigure}
	\begin{subfigure}[t]{.32\textwidth}
		\centering
		\includegraphics[width=.99\linewidth]{unob_stable}
		\caption{Selected robust Clustering}
		\label{fig:unob_stable}
	\end{subfigure}
	\begin{subfigure}[t]{.32\textwidth}
		\centering
		\includegraphics[width=.99\linewidth]{unob_consensus}
		\caption{Consensus Result}
		\label{fig:unob_consensus}
	\end{subfigure}	
	\caption{Single Clustering results for Data-Set}
	\label{fig:single}
	\end{figure}
\end{frame}

\begin{frame}{Tests: Multiple Solutions}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{multi}
		\caption{Example Data-Set with unknown Labels}
		\label{fig:multi}
	\end{figure}
\end{frame}

\begin{frame}{Tests: Multiple Solutions}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{multi_optics}
		\caption{OPTICS Plot}
		\label{fig:multi_optics}
	\end{figure}
\end{frame}

\begin{frame}{Tests: Multiple Solutions}
	\begin{figure}[h]
	\centering
	\begin{subfigure}[t]{.35\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{multi_c6}
		  \caption{Blue Cluster (19 base clu.)}
		  \label{fig:multi_c6}
	\end{subfigure}
	\begin{subfigure}[t]{.35\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{multi_c5_2}
		  \caption{Light-blue Cluster (14 base clu.)}
		  \label{fig:multi_c5_2}
	\end{subfigure}
	\begin{subfigure}[t]{.35\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{multi_c4}
		  \caption{Pink Cluster (9 base clu.)}
		  \label{fig:multi_c4}
	\end{subfigure}
	\begin{subfigure}[t]{.35\textwidth}
		  \centering
		  \includegraphics[width=.99\linewidth]{multi_c5_1}
		  \caption{Yellow Cluster (8 base clu.)}
		  \label{fig:multi_c5_1}
	\end{subfigure}

	\caption{Consensus Clustering results}
	\label{fig:multi_consres}
	\end{figure}
\end{frame}

\begin{frame}{Tests: User \& Real world data}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{user_wifi_gt}
		\caption{WiFi Localization Data-Set with first two Dimensions shown}
		\label{fig:user_wifi_gt}
	\end{figure}
\end{frame}

\begin{frame}{Tests: User \& Real world data}
	\begin{figure}[h]
		\centering
		\includegraphics[width=.64\textwidth]{user_wifi_optics}
		\caption{OPTICS Plot for WiFi Localization Data-Set}
		\label{fig:user_wifi_optics}
	\end{figure}
\end{frame}

\begin{frame}{Tests: User \& Real world data}

	\begin{figure}[h]
		\centering
		\begin{subfigure}[t]{.49\textwidth}
			  \centering
			  \includegraphics[width=.99\linewidth]{user_wifi_consensus}
			  \caption{User Consensus Result:\\NMI: 0.9142\\ blue (largest) Meta-Cluster}
			  \label{fig:user_wifi_consensus}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{.49\textwidth}
			  \centering
			  \includegraphics[width=.99\linewidth]{user_wifi_best}
			  \caption{Best Result of single Clustering Run\\NMI: 0.8904}
			  \label{fig:user_wifi_best}
		\end{subfigure}

		\caption{Clustering results for WiFi Localization Data-Set}
		\label{fig:user_wifi}
	\end{figure}
\end{frame}

\begin{frame}{Future Work}
Further evaluating usability:
	\begin{itemize}
	 	\item Additional study on usability
		\item Gathering information on which parts are especially useful
		\item Evaluating alternative views and functionality
		\item[]
	\end{itemize}

Research on consensus clustering:
	\begin{itemize}
	 	\item Analysis of generation/selection mechanisms
		\item Evaluation of selection criteria (is there a better choice than robustness)
	\end{itemize}

@@

	\begin{itemize}
		\item We created a new visual tool for cluster analysis:
		\begin{itemize}
		 	\item Visualizing clusterings on a meta-level
			\item Showing groups of robust clusterings
			\item Allowing to find solutions using consensus clustering
		\end{itemize}
		\item We showed:
		\begin{itemize}
		 	\item Robust groups indicate good results
		 	\item Combined results facilitate choice and can be better than any individual result
		\end{itemize}
		\item[]
		\item Link to the tool:
		\begin{itemize}
			\item \url{https://github.com/chris9182/Visual_Cluster_Exploration}
		\end{itemize}
	\end{itemize}

@@

\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}
